---
title: "software_carpentry_5"
date: "September 17, 2018"
output: html_document
---

#Adding columns in data frames
```{r}
age = c(2,3,5)
cats = read.csv(file="data/feline-data.csv")

cats = cbind(cats,age)
nrow(cats)
length(age)

```


#Adding rows in data frames (by adding a new level to a row of factors)
```{r}
newRow = list("tortoiseshell",3.3,TRUE,9)
cats = rbind(cats,newRow)
levels(cats$coat)
levels(cats$coat)=c(levels(cats$coat),"tortoiseshell")
cats = rbind(cats,newRow)
cats

```


#Changing factors into characters so you don't have to add new levels
```{r}
cats$coat = as.character(cats$coat)
str(cats)
```


#Challenge 1
```{r}
human_age = cats$age*7
human_age = as.factor(human_age)
human_age = as.numeric(human_age)
human_age = human_age/7

#When going from factor to numeric, it assigned each factor a numeric of 1, 2, 3, 4 instead of returning to its original numeric form
```

#Removing rows
```{r}
cats
cats[-4,] #same
na.omit(cats) #same
cats = na.omit(cats)
```

#Removing columns
```{r}
cats[,-4]

names(cats)

drop = names(cats) %in% c("age")
cats[,!drop]
```
#Removing columns the other way (doesn't really matter)
```{r}
drop = names(cats) %in% c("age")
cats[,!drop]
```


#Combining two data frames together
```{r}
cats = rbind(cats,cats)
cats

rownames(cats) = NULL
cats
```


#Challenge 2
#Do I have to do first_name = c("Paige") just for one answer? no
#Does it make a difference if I add the column with c or list? no
```{r}
my_info = data.frame(first_name = "Paige", last_name = "Bippus", lucky_number = 4, stringsAsFactors = FALSE)
my_info

alex_info = list("Alex","McCumber",7)
my_info = rbind(my_info, alex_info)
my_info

savannah_info = list("Savannah","Volkoff",8)
my_info = rbind(my_info, savannah_info)
my_info

answer = c("yes","yes","yes")
my_info = cbind(my_info, answer)
my_info
```

#Downloading and reading different types of files:
Files can also be downloaded directly from the Internet into a local folder of your choice onto your computer using the download.file function. The read.csv function can then be executed to read the downloaded file from the download location, for example,

download.file("https://raw.githubusercontent.com/swcarpentry/r-novice-gapminder/gh-pages/_episodes_rmd/data/gapminder_data.csv", destfile = "data/gapminder_data.csv")
gapminder <- read.csv("data/gapminder_data.csv")
Alternatively, you can also read in files directly into R from the Internet by replacing the file paths with a web address in read.csv. One should note that in doing this no local copy of the csv file is first saved onto your computer. For example,
gapminder <- read.csv("https://raw.githubusercontent.com/swcarpentry/r-novice-gapminder/gh-pages/_episodes_rmd/data/gapminder_data.csv")
You can read directly from excel spreadsheets without converting them to plain text first by using the readxl package.


#Realistic example
```{r}
gapminder = read.csv(file = "data/gapminder-FiveYearData.csv")

str(gapminder)

typeof(gapminder$year)
typeof(gapminder$country)
str(gapminder$country)

length(gapminder)
typeof(gapminder)

nrow(gapminder)
ncol(gapminder)
dim(gapminder)

colnames(gapminder)
head(gapminder)
```

#Challenge 3
```{r}
tail(gapminder)
gapminder[sample(nrow(gapminder), 5),]
```

#Challenge 4
```{r}
source(file = "scripts/gapminder_writein.R")
```

#Challenge 5
#What's the difference between an integer and number vector?
```{r}
str(gapminder)
```





