---
title: "software_carpentry_6"
output: html_document
---

#making a numeric vector
```{r}
x = c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) = c('a','b','c','d','e')
x
```

#accessing elements of a vector (subsetting a vector)
```{r}
x[1]
x[4]
x[c(3,5)]
x[1:4]
x[c(1,1,4)]
x[6]
```

#skipping and removing elements
```{r}
x[-2]
x[-c(2,5)]
x[-(1:3)] #can't do x[-1:3]
#can permanently remove a value from a vector by assigning it to x
```


#Challenge 1
```{r}
x = c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) = c('a','b','c','d','e')

x[-c(1,5)]
x[2:4]
```

#Subsetting by name
```{r}
x <- c(a=5.4, b=6.2, c=7.1, d=4.8, e=7.5) # we can name a vector 'on the fly'
x[c("a", "c")]
```

#Subsetting through other logical operators
```{r}
x[c(FALSE, FALSE, TRUE, FALSE, TRUE)]
x[x>7]
```

We often want to combine multiple logical criteria. For example, we might want to find all the countries that are located in Asia or Europe and have life expectancies within a certain range. Several operations for combining logical vectors exist in R:

&, the “logical AND” operator: returns TRUE if both the left and right are TRUE.
|, the “logical OR” operator: returns TRUE, if either the left or right (or both) are TRUE.
You may sometimes see && and || instead of & and |. These two-character operators only look at the first element of each vector and ignore the remaining elements. In general you should not use the two-character operators in data analysis; save them for programming, i.e. deciding whether to execute a statement.

!, the “logical NOT” operator: converts TRUE to FALSE and FALSE to TRUE. It can negate a single logical condition (eg !TRUE becomes FALSE), or a whole vector of conditions(eg !c(TRUE, FALSE) becomes c(FALSE, TRUE)).
Additionally, you can compare the elements within a single vector using the all function (which returns TRUE if every element of the vector is TRUE) and the any function (which returns TRUE if one or more elements of the vector are TRUE).

#Challenge 2
```{r}
x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) <- c('a', 'b', 'c', 'd', 'e')
print(x)

x[x>4 & x<7]
```

#Skipping named elements
```{r}
x <- c(a=5.4, b=6.2, c=7.1, d=4.8, e=7.5) # we start again by naming a vector 'on the fly'
x[-"a"]

x[names (x) !="a"]
```

#Recycling and excluding multiple values
```{r}
names(x) !=c("a","c") #because this vector is shorter than the x vector, R recycles c("a","c") over and over until it reaches the length of the first vector

x[! names(x) %in% c("a","c")]
```

#Challenge 3
```{r}
seAsia = c("Myanmar","Thailand","Cambodia","Vietnam","Laos")
gapminder = read.csv(file="data/gapminder-FiveYearData.csv", header=TRUE)
countries = unique(as.character(gapminder$country))

countries %in% seAsia
```

There are a number of special functions you can use to filter out this data:

is.na will return all positions in a vector, matrix, or data.frame containing NA (or NaN)
likewise, is.nan, and is.infinite will do the same for NaN and Inf.
is.finite will return all positions in a vector, matrix, or data.frame that do not contain NA, NaN or Inf.
na.omit will filter out all missing values from a vector

#Factor subsetting
```{r}
f = factor(c("a","a","b","c","c","d"))
f[f == "a"]
f[f %in% c("b","c")]
f[1:3]
f[-3]
```

#Matrix subsetting
#How is using set.seed and rnorm creating a matrix of random numbers?
```{r}
set.seed(1)
m = matrix(rnorm(6*4), ncol=4, nrow=6)
m

m[3:4, c(3,1)]
m[,c(3,4)]
m[3,] #One row or column automatically converted to a vector
m[3, , drop=FALSE] #Keeps the row or column as a matrix

m[5]

matrix(1:6, nrow=2, ncol=3, byrow=TRUE)
```

#Challenge 4
```{r}
m <- matrix(1:18, nrow=3, ncol=6)
print(m)

m[2,c(4,5)]
```

#List subsetting
```{r}
xlist = list(a="Software Carpentry", b=1:10, data=head(iris))
xlist
xlist[1]
xlist[1:2]
xlist[[1]] #Extracts an individual list of elements into a vector

xlist$data
```


#Challenge 5
```{r}
xlist = list(a="Software Carpentry", b=1:10, data=head(iris))
bvector = xlist[["b"]]
bvector[2]

xlist$b[2] #Easier solution
```

#Challenge 6
```{r}
mod <- aov(pop ~ lifeExp, data=gapminder)

mod
attributes(mod)
mod$df.residual
```








