---
title: "software_carpentry_4"
date: "September 7, 2018"
output: html_document
---

#Read in file
```{r}
data <- read.csv(file= "data/feline-data.csv")
data
```

```{r}
cats$weight
cats$coat
```

```{r}
cats$weight + 2
paste("My cat is", cats$coat)
```

#What type of data is a certain row or column?
```{r}
typeof(cats$weight)
typeof(3.14)
typeof(1L)
typeof(1)
typeof(1+1i)
typeof(TRUE)
typeof('banana')
```

#Added details of another cat
```{r}
cats <- read.csv(file= "data/feline-data_v2.csv")
cats
typeof(cats$weight)
class(cats)
```

#Reload file back to normal
```{r}
cats <- read.csv(file= "data/feline-data.csv")
cats
```


#dealing with vectors
```{r}
my_vector <- vector(length=3)
my_vector
another_vector <- vector(mode= 'character', length=3)
another_vector
str(cats$weight)

concat_vector <- c(2, 6, 3)
concat_vector
quiz_vector <- c(2, 6, '3')
quiz_vector

coercion_vector <- c('a', TRUE)
coercion_vector

another_coercion <- c(0, TRUE)
another_coercion
```

#coercing data to be a certain type
```{r}
character_vector_example <- c('0', '1', '2')
character_vector_example

character_coercedto_numeric <- as.numeric(character_vector_example)
character_coercedto_numeric

numeric_coercedto_logical <- as.logical(character_coercedto_numeric)
numeric_coercedto_logical

cats$likes_string
cats$likes_string <- as.logical(cats$likes_string)
cats$likes_string
```

#concatinating vectors and number series
```{r}
ab_vector <- c('a','b')
ab_vector

concat_example <- c(ab_vector, 'SWC')
concat_example

mySeries <- 1:10
mySeries

seq(10)
seq(1, 10, by=0.1)
```

#asking questions about vectors and giving names to elements
```{r}
sequence_example <- seq(10)
head(sequence_example, n=2)
tail(sequence_example, n=4)
length(sequence_example)
class(sequence_example)
typeof(sequence_example)

names_example <- 5:8
names(names_example) <- c("a","b","c","d")
names_example
```

#Challenge 1
```{r}
practice_vector <- 1:26
multiplied_vector <- practice_vector *2
names(multiplied_vector) <- LETTERS
multiplied_vector
```


#Data Frames
```{r}
str(cats$weight)
str(cats$likes_string)
str(cats$coat)
```

#Factors
```{r}
coats <- c('tabby','tortoiseshell','black','tabby')
coats
str(coats)
CATegories <- factor(coats)
str(CATegories)
typeof(coats)
typeof(CATegories)
```

#Challenge 2: issue with not pressing enter after end of feline-data.csv in order for R to read the table as complete
```{r}
cats <- read.csv(file= "data/feline-data.csv", stringsAsFactors=FALSE)
str(cats$coat)

```

#Ordering factors
```{r}
my_data <- c("case","control","control","case")
ordered_factors <- factor(my_data, levels= c("control","case"))
str(ordered_factors)
```

#Lists
```{r}
list_1 <- list(1, "a", TRUE, 1+4i)
list_1

list_2 <- list(title="Numbers", numbers=1:10, data=TRUE)
list_2

typeof(cats)

cats$coat
typeof(cats$coat)
cats[,1]
typeof(cats[,1])
str(cats[,1])

cats[1,]
typeof(cats[1,])
str(cats[1,])
```

#Challenge 3
```{r}
cats[1] #First column of the data frame (list)
cats[[1]] #Contents of the first column (a vector of factors)
cats$coat #Contents of the first column (a vector of factors)
cats["coat"] #First column of the data frame (list)
cats[1,1] #Value of row 1, column 1 (either factor or character)
cats[,1] #First column of the data frame (vector of factors)
cats[1,] #First row of the data frame (list)
```


#Matrices
```{r}
matrix_1 <- matrix(0, ncol=6, nrow=3)
matrix_1
class(matrix_1)
typeof(matrix_1)
str(matrix_1)
dim(matrix_1) #Dimensions
nrow(matrix_1)
ncol(matrix_1)
```

#Challenge 4
```{r}
length(matrix_1)
```

#Challenge 5
```{r}
matrix_2 <- matrix(1:50, ncol=5, nrow=10, byrow=TRUE)
matrix_2
```

#Challenge 6
```{r}
data_types <- c("double","integer","complex","logical","character")
data_structures <- c("data.frame","vector","factor","list","matrix")
answer <- list("data_types","data_structures")
answer
```

#Challenge 7
```{r}
matrix(c(4, 1, 9, 5, 10, 7), ncol=2, byrow=TRUE)
```


